---
title: "Data Visualizations for HRDP RNA Data from the Brain and Liver"
author: "Angela Yoder"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    fig_caption: yes
    fig_retina: 1 
    code_folding: hide
---

**The goals of this program are:**

1. Apply a PCA to the log2 transformed data after elimination of outliers
2. Determine the number of relevant PCs for adjustment
3. Export all PCs
4. Export residuals after adjustment by first 20 PCAs

```{r}
# Data Import

# set working directory
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data")

# download brain and liver datasets
brain = read.csv("PhenoGen.HRDP.v5.totalRNA.Brain.gene.ensembl96.txt", header = T, sep = "\t")
liver = read.csv("PhenoGen.HRDP.v5.totalRNA.Liver.gene.ensembl96.txt", header = T, sep = "\t")
```

```{r}

# Range of Genes

# distribution of gene ranges
range_brain = numeric(dim(brain)[1])
for (i in 1:dim(brain)[1])
{
  r = range(brain[i,])
  m = r[2] - r[1]
  range_brain[i] = 2^m
}

range_liver = numeric(dim(liver)[1])
for (i in 1:dim(liver)[1])
{
  r = range(liver[i,])
  m = r[2] - r[1]
  range_liver[i] = 2^m
}

hist(range_brain)
hist(range_liver)

```

```{r}

# Strain Means

# extracting strain from colname
rats_brain = sub("_.*", "", colnames(brain))
rats_liver = sub("_.*", "", colnames(liver))

# finding unique strains
strains_brain = unique(rats_brain)
strains_liver = unique(rats_liver)

# creating new data-frame : exchanging columns for rows
new_brain = t(brain)
new_liver = t(liver)

#get rid of ERCC cols

  #indices of cols with ERCC
elim_liver = grep("ERCC", colnames(new_liver))
elim_brain = grep("ERCC", colnames(new_brain))
  #creating new end point for columns
newdim_liver = elim_liver[1] - 1
newdim_brain = elim_brain[1] -1
  #cutting off ERCC cols
new_brain = new_brain[,1:newdim_brain]
new_liver = new_liver[,1:newdim_liver]

new_brain = as.data.frame(new_brain)
new_liver = as.data.frame(new_liver)

# vector for gene names
genes_brain = colnames(new_brain[,2:dim(new_brain)[2]])
genes_liver = colnames(new_liver[,2:dim(new_liver)[2]])

# adding a strains as factor column to new matrix
strain_factor_brain = rats_brain
strain_factor_liver = rats_liver

new_brain = cbind(strain_factor_brain, new_brain)
new_liver = cbind(strain_factor_liver, new_liver)

new_brain$strain_factor_brain = factor(new_brain$strain_factor_brain, levels = strains_brain)
new_liver$strain_factor_liver = factor(new_liver$strain_factor_liver, levels = strains_liver)

# get number of genes for each dataset
num_col_brain = length(genes_brain) 
num_col_liver = length(genes_liver) 

# creating a matrix to store means strainXgene
storagematrix_brain = matrix(NA, nrow = length(strains_brain), ncol = num_col_brain, dimnames = list(strains_brain, genes_brain)) 
storagematrix_liver = matrix(NA, nrow = length(strains_liver), ncol = num_col_liver, dimnames = list(strains_liver, genes_liver))

# calculate the means within each strain within each gene
for (i in 1:length(strains_brain)) # going through each strain
{
  b = new_brain[new_brain$strain_factor_brain == strains_brain[i],] # extracting the rows where strain matches
  for (j in 2:dim(new_brain)[2]) # Going through each gene
  {
    v = mean(as.numeric(b[,j])) # calculating mean across the strain for each gene
    r = j-1
    v = format(round(v, 2), nsmall = 2)
    v = as.numeric(v)
    storagematrix_brain[i, r] = v # storing
  }
}

for (i in 1:length(strains_liver)) # going through each strain
{
  b = new_liver[new_liver$strain_factor_liver == strains_liver[i],] # extracting the rows where strain matches
  for (j in 2:dim(new_liver)[2]) # Going through each gene
  {
    v = mean(as.numeric(b[,j])) # calculating mean across the strain for each gene
    r = j-1
    v = format(round(v, 2), nsmall = 2)
    v = as.numeric(v)
    storagematrix_liver[i, r] = v # storing
  }
}

# RLE


median_brain = as.numeric(apply(storagematrix_brain, MARGIN = 2, FUN = median))
median_liver = as.numeric(apply(storagematrix_liver, MARGIN = 2, FUN = median))


RLE_brain = matrix(NA, nrow = dim(storagematrix_brain)[1], ncol = dim(storagematrix_brain)[2])
for (i in 1:length(storagematrix_brain[1,]))
{
  RLE_brain[,i] = abs(median_brain[i] - storagematrix_brain[,i])
}

RLE_liver = matrix(NA, nrow = dim(storagematrix_liver)[1], ncol = dim(storagematrix_liver)[2])
for (i in 1:length(storagematrix_liver[1,]))
{
  RLE_liver[,i] = abs(median_liver[i] - storagematrix_liver[,i])
}

colnames(RLE_brain) = colnames(storagematrix_brain)
colnames(RLE_liver) = colnames(storagematrix_liver)
rownames(RLE_brain) = rownames(storagematrix_brain)
rownames(RLE_liver) = rownames(storagematrix_liver)


boxplot(x = t(RLE_brain), main = "RLE plot: Brain dataset", xlab = "Strain")
boxplot(x = t(RLE_liver), main = "RLE plot: Liver dataset", xlab = "Strain")

```

```{r}

#Broad Sense Heritability

library(broom)

formula_brain = lapply(colnames(new_brain[,2:length(new_brain[1,])]), function (x) as.formula(paste(x, " ~ strain_factor_brain")))
formula_liver = lapply(colnames(new_liver[,2:length(new_liver[1,])]), function (x) as.formula(paste(x, " ~ strain_factor_liver")))

BSH_brain = lapply(formula_brain, function(x) tidy(aov(x, data = new_brain))$sumsq)
BSH_liver = lapply(formula_liver, function(x) tidy(aov(x, data = new_liver))$sumsq)

r.squared_brain = lapply(BSH_brain, function(a) a[1]/(a[1] + a[2]))
r.squared_liver = lapply(BSH_liver, function(a) a[1]/(a[1] + a[2]))
r.squared_brain = unlist(r.squared_brain)
r.squared_liver = unlist(r.squared_liver)
names_brain = as.character(formula_brain)
names_liver = as.character(formula_liver)

fin.r.squared_brain = cbind(names_brain, r.squared_brain)
fin.r.squared_liver = cbind(names_liver, r.squared_liver)

hist(as.numeric(fin.r.squared_brain[,2]))
hist(as.numeric(fin.r.squared_liver[,2]))
```

```{r}

#Hierarchical Clustering

for_clustering_brain = as.dist(1 - abs(cor(x = brain, method = "spearman")))
for_clustering_liver = as.dist(1 - abs(cor(x = liver, method = "spearman")))

h_clustering_brain = hclust(for_clustering_brain)
h_clustering_liver = hclust(for_clustering_liver)

plot(h_clustering_brain)
plot(h_clustering_liver)

```

Different clustering strategies?

```{r}

# PCA Plot

# PCA
library(PCAtools)

# extracting strain from colname
rats_brain = sub("_.*", "", colnames(brain))
rats_liver = sub("_.*", "", colnames(liver))

pca_brain <- pca(brain, center=TRUE, scale=TRUE)
pca_liver = pca(liver, center = TRUE, scale = TRUE)

biplot(pca_brain)
biplot(pca_liver)
