---
title: "WGCNA Section 4: Interfacing network analysis with other data such as functional
annotation and gene ontology"
author: "Angela Yoder"
date: "`r Sys.Date()`"
---

**The goals of this program are:**

1. Find problems with enrichmentAnalysis

It seems there are 1115 rows with a significant value for FDR (less than 0.05). Setting the threshold type seeems to change some element of row selection as total number of rows decrease, but does not map one-to-one onto the FDR values less than 0.05. The only time the number of FDR values less than 0.01 decreases is if the threshold is very low for FDR, using a stricter analysis like Bonferroni
Check if the 1115 are the same for each iteration. The dataframes produced by running the enrichmentAnalysis and then pulling out values of FDR < 0.05 appear to be the same. 

```{r, echo = F, results = 'hide', message = F, warning = F}
# Data input

# set working directory
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")

# upload data
phen = read.csv("Phenotype_WGCNA.csv")
protein_brain = read.csv("Brain_WGCNA.csv")
moduleColors= read.csv("moduleColors_WGCNA.csv")
chosenModules = read.csv("Chosen Modules.csv")

# fix problems
moduleColors = moduleColors[,2]
names = protein_brain[,1]
protein_brain = protein_brain[,2:dim(protein_brain)[2]]
rownames(protein_brain) = names

names = phen[,1]
phen = phen[,2:5]
rownames(phen) = names 

# Read in the probe annotation
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
annot = read.csv(file = "Annotation.csv");
annot = annot[,2:dim(annot)[2]]

library(WGCNA)
library(tidyverse)


# Match probes in the data set to the probe IDs in the annotation file
probes = names(protein_brain)
probes2annot = match(probes, annot$ensembl_gene_id)
# Get the corresponding Locuis Link IDs
allLLIDs = annot$external_gene_name[probes2annot];



```

FDR threshold 0.05
```{r}
library(anRichment)
GOcollection = buildGOcollection(organism = "rat")

entrez = convert2entrez(organism = "rat", symbol = allLLIDs);

#GOenr = enrichmentAnalysis(classLabels = moduleColors, identifiers = allLLIDs);

#GOenr = GOenrichmentAnalysis(moduleColors, allLLIDs, organism = "rat", nBestP = 10);
GOenr = enrichmentAnalysis(
  classLabels = moduleColors, identifiers = entrez,
  refCollection = GOcollection,
  useBackground = "given",
  threshold = 0.05,
  getFDR = T, 
  thresholdType = "FDR",
  getOverlapEntrez = TRUE,
  getOverlapSymbols = TRUE,
  ignoreLabels = "grey");



tab = GOenr$enrichmentTable




```
Bonferroni 0.05
```{r}

GOenr = enrichmentAnalysis(
  classLabels = moduleColors, identifiers = entrez,
  refCollection = GOcollection,
  useBackground = "given",
  threshold = 0.05,
  getFDR = T, 
  thresholdType = "Bonferroni",
  getOverlapEntrez = TRUE,
  getOverlapSymbols = TRUE,
  ignoreLabels = "grey");



bontab = GOenr$enrichmentTable

```

Nominal 0.05
```{r}

GOenr = enrichmentAnalysis(
  classLabels = moduleColors, identifiers = entrez,
  refCollection = GOcollection,
  useBackground = "given",
  threshold = 0.05,
  getFDR = T, 
  thresholdType = "nominal",
  getOverlapEntrez = TRUE,
  getOverlapSymbols = TRUE,
  ignoreLabels = "grey");



nom5tab = GOenr$enrichmentTable

```

```{r}

## look at total number

dim(tab)
dim(bontab)
dim(nom5tab)

## look at number below threshold

dim(tab[tab$FDR < 0.05,])
dim(bontab[bontab$FDR < 0.05,])
dim(nom5tab[nom5tab$FDR < 0.05,])

```

FDR 0.10
```{r}

GOenr = enrichmentAnalysis(
  classLabels = moduleColors, identifiers = entrez,
  refCollection = GOcollection,
  useBackground = "given",
  threshold = 0.10,
  getFDR = T, 
  thresholdType = "FDR",
  getOverlapEntrez = TRUE,
  getOverlapSymbols = TRUE,
  ignoreLabels = "grey");



FDR10tab = GOenr$enrichmentTable

```

FDR 0.001
```{r}

GOenr = enrichmentAnalysis(
  classLabels = moduleColors, identifiers = entrez,
  refCollection = GOcollection,
  useBackground = "given",
  threshold = 0.000001,
  getFDR = T, 
  thresholdType = "nominal",
  getOverlapEntrez = TRUE,
  getOverlapSymbols = TRUE,
  ignoreLabels = "grey");



FDR001tab = GOenr$enrichmentTable

```

```{r}

dim(tab)
dim(FDR10tab)
dim(FDR01tab)

dim(tab[tab$FDR < 0.05,])
dim(FDR10tab[FDR10tab$FDR < 0.05,])
dim(FDR01tab[FDR01tab$FDR < 0.05,])

table(tab[tab$FDR < 0.05, ] == nom5tab[nom5tab$FDR < 0.05, ], useNA = 'ifany')
table(tab[tab$FDR < 0.05, ] == nom10tab[nom10tab$FDR < 0.05, ], useNA = 'ifany')

bee = tab[tab$FDR < 0.05, ]
gee = nom5tab[nom5tab$FDR < 0.05, ]
```