```{r}

# set working directory
setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")

# phenotypes dataset
phen = read.csv("phenotypesAcetate.csv")
phen_strain = phen[,2]
phen = phen[,3:4]
rownames(phen) = phen_strain

# download strain means dataset
strainmeans_brain = read.csv("Means_strainXgene_brain.csv")
#strainmeans_liver = read.csv("Means_strainXgene_liver.csv")
rowname_b = strainmeans_brain[,1]
strainmeans_brain = strainmeans_brain[,2:dim(strainmeans_brain)[2]]
rownames(strainmeans_brain) = rowname_b

# only RI

hxb_brain = grep("HXB", rownames(strainmeans_brain))
bxh_brain = grep("BXH", rownames(strainmeans_brain))
recombinant_brain = sort(c(hxb_brain,bxh_brain))
recomblog_brain = 1:length(rownames(strainmeans_brain))
recomblog_brain[recombinant_brain] = 0
recomblog_brain = recomblog_brain == 0

strainmeans_brain = strainmeans_brain[recomblog_brain,]

# libraries
library(WGCNA)
library(tidyverse)
library(biomaRt)

# getting only protein-coding genes from our dataset
ensembl = useEnsembl(biomart="ensembl")
ensembl = useEnsembl(biomart="ensembl",dataset = "rnorvegicus_gene_ensembl")
get_anno = getBM(attributes=c("ensembl_gene_id",
                              "external_gene_name",
                              "description",
                              "chromosome_name",
                              "start_position",
                              "end_position",
                              "gene_biotype"),
                 filters = 'biotype',
                 values = "protein_coding",
                 mart = ensembl)

findpro_brain = match(get_anno$ensembl_gene_id, colnames(strainmeans_brain))
findpro_brain = findpro_brain %>%
  na.omit %>%
  sort


protein_brain = strainmeans_brain[,findpro_brain]

order_strain = c()
for (i in rownames(protein_brain))
{
  order_strain = c(order_strain, grep(paste(i, "$", sep = ""), rownames(phen)))
}

phen = phen[order_strain,]

#exp_protein_brain = 2 ^ protein_brain 

```

In this tutorial section, we illustrate the 1-step, automatic multiple set network construction and detection of
consensus modules. We note that while the actual network construction and module detection is executed in a
single function call, a preliminary step of choosing a suitable soft-thresholding power must be performed first.

Uncertain: choosing soft-thresholding powers, choosing module size [size 5]. blockwiseConsensusmodules has many parameters- may need to be adjusted; deepsplit- 4
```{r}

# Choose a set of soft-thresholding powers

powers = c(c(1:10), seq(from = 12, to=20, by=2))
# Call the network topology analysis function
sft = pickSoftThreshold(protein_brain, powerVector = powers, verbose = 5) ## expression data
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h; use 0.8 to 0.95
abline(h=0.80,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")

```

plotdendroandcolors is module eigengenes?
```{r}
softPower = 8
adjacency = adjacency(protein_brain, power = softPower)

TOM = TOMsimilarity(adjacency);
dissTOM = 1 - TOM

geneTree = hclust(as.dist(dissTOM), method = "average")
plot(geneTree, labels = FALSE)

# We like large modules, so we set the minimum module size relatively high:
minModuleSize = 30;
# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
deepSplit = 2, pamRespectsDendro = FALSE,
minClusterSize = minModuleSize);
table(dynamicMods)

# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05,
main = "Gene dendrogram and module colors")

#### OTHER ####

bwnet = blockwiseModules(protein_brain, maxBlockSize = 5000,
power = 8, TOMType = "unsigned", minModuleSize = 30,
reassignThreshold = 0, mergeCutHeight = 0.25,
numericLabels = TRUE,
saveTOMs = TRUE,
saveTOMFileBase = "TOM-blockwise",
verbose = 3)

bwLabels = matchLabels(bwnet$colors, moduleLabels);
# Convert labels to colors for plotting
bwModuleColors = labels2colors(bwLabels)

plotDendroAndColors(bwnet$dendrograms[[2]], bwModuleColors[bwnet$blockGenes[[2]]],
"Module colors", main = "Gene dendrogram and module colors in block 1",
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)

sizeGrWindow(12,9)
plotDendroAndColors(geneTree,
cbind(moduleColors, bwModuleColors),
c("Single block", "2 blocks"),
main = "Single block gene dendrogram and module colors",
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)

```


```{r}
# Calculate eigengenes
MEList = moduleEigengenes(protein_brain, colors = dynamicColors)
MEs = MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);
# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
plot(METree, main = "Clustering of module eigengenes",
xlab = "", sub = "")

MEDissThres = 0.25
# Plot the cut line into the dendrogram
abline(h=MEDissThres, col = "red")
# Call an automatic merging function
merge = mergeCloseModules(protein_brain, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors;
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs;
sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
c("Dynamic Tree Cut", "Merged dynamic"),
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
#dev.off()

# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs

```

## sec 3

Show association between module eigengenes and clinical traits

```{r}

# Define numbers of genes and samples
nGenes = ncol(protein_brain);
nSamples = nrow(protein_brain);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(protein_brain, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, phen, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)


sizeGrWindow(10,6)
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
cex.lab.y = 0.35,
xLabels = names(phen),
yLabels = names(MEs),
ySymbols = names(MEs),
colorLabels = FALSE,
colors = greenWhiteRed(50),
textMatrix = textMatrix,
setStdMargins = FALSE,
cex.text = 0.25,
zlim = c(-1,1),
main = paste("Module-trait relationships"))



# Define variable weight containing the weight column of datTrait
auc1 = as.data.frame(phen$meanauc1gkg);
names(auc1) = "auc1"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(protein_brain, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(protein_brain, auc1, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(auc1), sep="");
names(GSPvalue) = paste("p.GS.", names(auc1), sep="")

##

for (i in 1: dim(moduleTraitCor)[1])
{
  # pick out max and min for each column bleh  
}
```

example of looking at a gene and its association to a trait
```{r}

module = "brown"
column = match(module, modNames);
moduleGenes = moduleColors==module;
sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
xlab = paste("Module Membership in", module, "module"),
ylab = "Gene significance for auc1",
main = paste("Module membership vs. gene significance\n"),
cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

```

Next part has to do with identifying gene names- not sure i can do at this step

## Section 4
Also requires gene identification data

```{r}
# Read in the probe annotation
annot = read.csv(file = "GeneAnnotation.csv");
# Match probes in the data set to the probe IDs in the annotation file
probes = names(datExpr)
probes2annot = match(probes, annot$substanceBXH)
# Get the corresponding Locuis Link IDs
allLLIDs = annot$LocusLinkID[probes2annot];
# $ Choose interesting modules
intModules = c("brown", "red", "salmon")
for (module in intModules)
{
# Select module probes
modGenes = (moduleColors==module)
# Get their entrez ID codes
modLLIDs = allLLIDs[modGenes];
# Write them into a file
fileName = paste("LocusLinkIDs-", module, ".txt", sep="");
write.table(as.data.frame(modLLIDs), file = fileName,
row.names = FALSE, col.names = FALSE)
}
# As background in the enrichment analysis, we will use all probes in the analysis.
fileName = paste("LocusLinkIDs-all.txt", sep="");
write.table(as.data.frame(allLLIDs), file = fileName,
row.names = FALSE, col.names = FALSE)

```

## Section 5

```{r}

nGenes = ncol(protein_brain)
nSamples = nrow(protein_brain)

# Calculate topological overlap anew: this could be done more efficiently by saving the TOM
# calculated during module detection, but let us do it again here.
dissTOM = 1-TOMsimilarityFromExpr(protein_brain, power = 6);
# Transform dissTOM with a power to make moderately strong connections more visible in the heatmap
plotTOM = dissTOM^7;
# Set diagonal to NA for a nicer plot
diag(plotTOM) = NA;
# Call the plot function
sizeGrWindow(9,9)
TOMplot(plotTOM, geneTree, moduleColors, main = "Network heatmap plot, all genes")

nSelect = 400
# For reproducibility, we set the random seed
set.seed(10);
select = sample(nGenes, size = nSelect);
selectTOM = dissTOM[select, select];
# There’s no simple way of restricting a clustering tree to a subset of genes, so we must re-cluster.
selectTree = hclust(as.dist(selectTOM), method = "average")
selectColors = moduleColors[select];
# Open a graphical window
sizeGrWindow(9,9)
# Taking the dissimilarity to a power, say 10, makes the plot more informative by effectively changing
# the color palette; setting the diagonal to NA also improves the clarity of the plot
plotDiss = selectTOM^7;
diag(plotDiss) = NA;
TOMplot(plotDiss, selectTree, selectColors, main = "Network heatmap plot, selected genes")


# Recalculate module eigengenes
MEs = moduleEigengenes(protein_brain, moduleColors)$eigengenes
# Isolate weight from the clinical traits
weight = as.data.frame(phen[,1]);
names(weight) = "weight"
# Add the weight to existing module eigengenes
MET = orderMEs(cbind(MEs, weight))
# Plot the relationships among the eigengenes and the trait
sizeGrWindow(5,7.5);
par(cex = 0.9)
plotEigengeneNetworks(MET, "", marDendro = c(0,4,1,2), marHeatmap = c(3,4,1,2), cex.lab = 0.8, xLabelsAngle
= 90)

# Plot the dendrogram
sizeGrWindow(6,6);
par(cex = 1.0)
plotEigengeneNetworks(MET, "Eigengene dendrogram", marDendro = c(0,4,2,0),
plotHeatmaps = FALSE)
# Plot the heatmap matrix (note: this plot will overwrite the dendrogram plot)
par(cex = 1.0)
plotEigengeneNetworks(MET, "Eigengene adjacency heatmap", marHeatmap = c(3,4,2,2),
plotDendrograms = FALSE, xLabelsAngle = 90)

```